### 2019-02-23 doc/model_descibe.tex
updating graphs. everything runs headless. table of model desc

\documentclass[12pt,letterpaper]{article}

\usepackage[margin=1cm]{geometry}

\begin{document}

\begin{tabular}[ht]{ l l l l }
  Code & Description & Covariates & Formula \\
  \hline
  C & Constant effects, no historical covariates & Geographic range, global temperature & $ event \sim range + temp + (1 | age/phylum) $ \\
  V & Varying effects, no historical covariates & Geographic range, global temperature  & $ event \sim range + temp + (1 + range + temp | mybin/phylum) + (1 | age/phylum) $ \\
  CP & Constant effects, historical covariates & Geographic range, change in geographic range, global temperature, previous temperature & $ event \sim range + range\_diff + temp + temp\_lag + (1 | age/phylum) $ \\
  VP & Varying effects, historical covariates & Geographic range, change in geographic range, global temperature, previous temperature & $ event \sim range + range\_diff + temp + temp\_lag + (1 + range + range\_diff + temp + temp\_lag | mybin/phylum) + (1 | age/phylum) $ \\
\end{tabular}

{
  \def\arraystretch{1.5}
  \begin{tabular}[ht]{ l p{3cm} l l }
    Code & Description & Covariates & Formula \\
    \hline
    C & Constant effects, no historical cov. & \parbox[t]{0.25\textwidth}{Geographic range,\\temperature} & \parbox[t]{0.33\textwidth}{event $\sim$ range + temp +\\(1 $|$ age/phylum)} \\
    V & Varying effects, no historical cov. & \parbox[t]{0.25\textwidth}{Geographic range,\\temperature} & \parbox[t]{0.33\textwidth}{event $\sim$ range + temp +\\(1 + range + temp $|$ phylum) +\\(1 $|$ age/phylum)} \\ 
    CP & Constant effects, historical cov. & \parbox[t]{0.25\textwidth}{Geographic range,\\change in geographic range,\\temperature,\\previous temperature} & \parbox[t]{0.33\textwidth}{event $\sim$ range + range\_diff +\\temp + temp\_lag +\\(1 $|$ age/phylum)} \\
    VP & Varying effects, historical cov. & \parbox[t]{0.25\textwidth}{Geographic range,\\change in geographic range,\\temperature,\\previous temperature} & \parbox[t]{0.33\textwidth}{event $\sim$ range + range\_diff +\\temp + temp\_lag +\\(1 + range + range\_diff + temp + temp\_lag $|$ phylum) +\\(1 $|$ age/phylum)} \\
  \end{tabular}
}

\end{document}