### 2019-02-25 doc/model_descibe.tex
seth comments on manuscript filtered through. supp info. fewer figures. regen figures.

\documentclass[12pt,letterpaper]{article}

\usepackage[margin=1cm]{geometry}

\begin{document}

{
  \def\arraystretch{1.5}
  \begin{tabular}[ht]{ l p{3cm} l l }
    Code & Description & Covariates & Formula \\
    \hline
    C & Constant effects, no historical cov. & \parbox[t]{0.25\textwidth}{Geographic range,\\temperature} & \parbox[t]{0.33\textwidth}{event $\sim$ range + temp +\\(1 $|$ age/phylum)} \\
    V & Varying effects, no historical cov. & \parbox[t]{0.25\textwidth}{Geographic range,\\temperature} & \parbox[t]{0.33\textwidth}{event $\sim$ range + temp +\\(1 + range + temp $|$ phylum) +\\(1 $|$ age/phylum)} \\ 
    CP & Constant effects, historical cov. & \parbox[t]{0.25\textwidth}{Geographic range,\\change in geographic range,\\temperature,\\previous temperature} & \parbox[t]{0.33\textwidth}{event $\sim$ range + range\_diff +\\temp + temp\_lag +\\(1 $|$ age/phylum)} \\
    VP & Varying effects, historical cov. & \parbox[t]{0.25\textwidth}{Geographic range,\\change in geographic range,\\temperature,\\previous temperature} & \parbox[t]{0.33\textwidth}{event $\sim$ range + range\_diff +\\temp + temp\_lag +\\(1 + range + range\_diff + temp + temp\_lag $|$ phylum) +\\(1 $|$ age/phylum)} \\
  \end{tabular}
}

\end{document}